Visual Diagnostics for Constrained Optimisation with Application to Guided Tours

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.Rmd
a missing action
introduction modified

The focus of this paper is on the optimisation problem arising in the projection pursuit guided tour [@buja2005computational], an exploratory data analysis tool used for detecting interesting structures in high-dimensional data through a set of lower-dimensional projections [@cook2008grand]. The goal of the optimisation is to identify the projection, represented by the projection matrix, that gives the most interesting low-dimensional view. A view is interesting if it can show some structures of the data other than a normal distribution, e.g., bi-modality, clustering, and outliers.

The optimization challenges encountered in the projection pursuit guided tour problem are common to those of optimization in general. Examples include the existence of multiple optima (local and global), the trade-off between computational burden and proximity to the optima, or dealing with noisy objective functions that might be non-smooth and non-differentiable [@jones1998efficient]. The visualization tools and optimization methods presented in this paper can therefore easily be applied to other optimization problems.

The remainder of the paper is organised as follows. 
Section \ref{optim} provides an overview of optimisation methods, specifically line search methods. 
Section \ref{tour} reviews projection pursuit guided tour, defines the optimisation problem and outlines three existing algorithms.
Section \ref{vis-diag} presents the new visual diagnostics. A data structure is defined to capture information during the optimisation, and different diagnostic plots are designed with the data collected.
Section \ref{application} shows applications of how these diagnostic plots can be used to examine the performance of different optimisers and guide improvements to the existing algorithms.
Finally, Section \ref{implementation} describes the R package: \CRANpkg{ferrn} [@ferrn], that implements the visual diagnostics.

>>>>

The focus of this paper is on the optimisation problem arising in the projection pursuit guided tour [@buja2005computational], an exploratory data analysis technique used for detecting interesting structures in high-dimensional data through a set of lower-dimensional projections [@cook2008grand]. The goal of the optimisation is to identify the projection, represented by the projection matrix, that gives the most interesting low-dimensional view. A view is interesting if it can show some structures of the data that depart from normalality, such as bimodality, clustering, or outliers.

The optimization challenges encountered in the projection pursuit guided tour problem are common to those of optimization in general. Examples include the existence of multiple optima (local and global), the trade-off between computational burden and proximity to the optima, or dealing with noisy objective functions that might be non-smooth and non-differentiable [@jones1998efficient]. The visualization tools, optimization methods and conceptual framework presented in this paper can therefore be applied to other optimization problems.

The remainder of the paper is organised as follows. 
The next section provides an overview of optimisation methods, specifically line search methods.  A review of the projection pursuit guided tour, an overview of the optimisation problem and outlines of three existing algorithms follows.
The third section presents the new visual diagnostics, including the the design of a data structure to capture information during the optimisation, from which several diagnostic plots are created.
An illustration of how the diagnostic plots can be used to examine the performance of different optimisers and guide improvements to the existing algorithms applications is shown using simulated data.
Finally, an explanation of the \ref{implementation} in the R package, \CRANpkg{ferrn} [@ferrn], is provided.

### 2021-03-25 paper/zhang-cook-laa-langrene-menendez.tex
introduction modified

The focus of this paper is on the optimisation problem arising in the
projection pursuit guided tour \citep{buja2005computational}, an
exploratory data analysis tool used for detecting interesting structures
in high-dimensional data through a set of lower-dimensional projections
\citep{cook2008grand}. The goal of the optimisation is to identify the
projection, represented by the projection matrix, that gives the most
interesting low-dimensional view. A view is interesting if it can show
some structures of the data other than a normal distribution, e.g.,
bi-modality, clustering, and outliers.

The optimization challenges encountered in the projection pursuit guided
tour problem are common to those of optimization in general. Examples
include the existence of multiple optima (local and global), the
trade-off between computational burden and proximity to the optima, or
dealing with noisy objective functions that might be non-smooth and
non-differentiable \citep{jones1998efficient}. The visualization tools
and optimization methods presented in this paper can therefore easily be
applied to other optimization problems.

The remainder of the paper is organised as follows. Section \ref{optim}
provides an overview of optimisation methods, specifically line search
methods. Section \ref{tour} reviews projection pursuit guided tour,
defines the optimisation problem and outlines three existing algorithms.
Section \ref{vis-diag} presents the new visual diagnostics. A data
structure is defined to capture information during the optimisation, and
different diagnostic plots are designed with the data collected. Section
\ref{application} shows applications of how these diagnostic plots can
be used to examine the performance of different optimisers and guide
improvements to the existing algorithms. Finally, Section
\ref{implementation} describes the R package: \CRANpkg{ferrn}
\citep{ferrn}, that implements the visual diagnostics.

>>>>

The focus of this paper is on the optimisation problem arising in the
projection pursuit guided tour \citep{buja2005computational}, an
exploratory data analysis technique used for detecting interesting
structures in high-dimensional data through a set of lower-dimensional
projections \citep{cook2008grand}. The goal of the optimisation is to
identify the projection, represented by the projection matrix, that
gives the most interesting low-dimensional view. A view is interesting
if it can show some structures of the data that depart from normalality,
such as bimodality, clustering, or outliers.

The optimization challenges encountered in the projection pursuit guided
tour problem are common to those of optimization in general. Examples
include the existence of multiple optima (local and global), the
trade-off between computational burden and proximity to the optima, or
dealing with noisy objective functions that might be non-smooth and
non-differentiable \citep{jones1998efficient}. The visualization tools,
optimization methods and conceptual framework presented in this paper
can therefore be applied to other optimization problems.

The remainder of the paper is organised as follows. The next section
provides an overview of optimisation methods, specifically line search
methods. A review of the projection pursuit guided tour, an overview of
the optimisation problem and outlines of three existing algorithms
follows. The third section presents the new visual diagnostics,
including the the design of a data structure to capture information
during the optimisation, from which several diagnostic plots are
created. An illustration of how the diagnostic plots can be used to
examine the performance of different optimisers and guide improvements
to the existing algorithms applications is shown using simulated data.
Finally, an explanation of the \ref{implementation} in the R package,
\CRANpkg{ferrn} \citep{ferrn}, is provided.
